---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Alexander Alexandrov"
output: pdf_document
mode: selfcontained
---

# Overview

This paper contains some basic analysis of the ToothGrowth data.
Analysis object is effectivness of different supplement types against teeth response.
The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

# Tooth Growth Data Overview

Load the tooth growth data.

```{r}
library(datasets)
library(data.table)
data <- data.table(ToothGrowth)
dose.levels <- as.numeric(levels(factor(data$dose)))
str(data)
```

Data is clean anough.

|Variable|Type|Domain|Description|
|---|---|---|---|
|**len**|numeric|positive real number|Tooth length|
|**supp**|factor|string: "VC" (ascorbic acid), "OJ" (orange juice)|Supplement type|
|**len**|numeric|positive real number: 0.5, 1.0, 2.0|Dose in milligrams|

\newpage

# Further exploratory data analyses

Some usefull plots.

```{r}
library(ggplot2)
qplot(dose, len, data = data, size = 2, colour = supp, geom = c("boxplot", "point"),
      alpha = 0.2, main = "Response by supplement type", ylab = "Response (length)")
```

According to quantiles (boxplots) orange juice (red) is slightly better then ascorbic acid.

\newpage

Next group response by dose and supplement type.

```{r}
qplot(supp, len, data = data, size = 2, colour = supp, geom = c("boxplot", "point"),
      alpha = 0.2, facets = ~dose,
      main = "Response by dose and supplement type", ylab = "Response (length)")
```

It's evident from figure that orange juice is preferable only in case of the small doses.
Results for the 2.0 mg dose are very close to each other.

\newpage

# Confidence intervals

Paired confidence intervals (orange juice vs ascorbic acid) for each dose.

```{r}
ComputeConfInt <- function(d) {
    conf.int <- t.test(len ~ supp, data = data[dose == d], paired = F)$conf.int
    return(data.table(dose = d, low = conf.int[1], high = conf.int[2]))
}
conf.intervals <- rbindlist(lapply(dose.levels, ComputeConfInt))
conf.intervals
```

So with the 95% confidence following assumptions can be made:

|Dose (mg)|OJ-VC teeth response 95% conf. int.|Description|
|---|---|---|
|**0.5**|from 1.7 to 8.8|Conf. int. is positive, so orange juice is the best choise for small doses|
|**1.0**|from 2.8 to 9.1||
|**2.0**|from -3.8 to 3.6|Orange juice has slightly same effect compared to ascorbic acid|
